\documentclass{sagej}
\usepackage{amsmath}
\usepackage{url}
\begin{document}
\title{Calculating atmospheric IR absorbtion}
\author{Dr Barry D.O. Adams\affilnum{1}}
\affiliation{This work was self funded, Dr Adams is employed in a Back End Developer Role not related to this work}
\corrauth{}
\email{barry.david.adams@gmail.com}
\begin{abstract}
  In this work, we present a program to numerical integrate the atmospheric absorbtion due to carbon dioxide and water vapour
	of infrared photons by the atmosphere.

\end{abstract}
\keywords{Global Warming, Carbon Dioxide, Water Vapour}
\maketitle
\section{Introduction}
   In recent years, the hypothosis that human released carbon dioxide will warm planet earth has become an huge issue. There is naturally a
large body of work in it. However very few of the calculation tools for the issue are released as open source. In this work we present
a program and it results to numerical integrate carbon dioxide and water vapour absorption over frequency, latitude and height, producing a percentage
absorbtion ammount for the two green house gases.

\section{Input Absorption Lines}
Spectra absorbtion lines for many molecules are available on the HITRAN database online. We are interest in the strongest lines for absorbtion by the
molecules with the most common isotopes from the ground state, since it is the ground state is far more common than the higher levels. Download the HITRAN
frequency, upper and lower quantum states, with the Einstein $A$ Coefficent that gives the absorption strength of the lines. Using the line intersity $S$, from
HITRAN we order the lines from strongest to weakest, and select those starting at ground strength. We find 3 lines for carbon dioxide and 4 for water, and present them in the
table below. The $CO_2$ lines reference is given by HITRAN is \cite{Zak}, and for water vapour it is \cite{},

\begin{table}
	\caption{The most prominent absorption lines from ground state}
	\centering
	\begin{tabular} { c c c c c c c}
        \hline\hline
		Molecule & Wave-number $(cm^-1)$ & Einstein Coefficient $A$ & upper state multiplicity & upper state & line width air at $1 atm$ & $n_{air}$ \\ 
		\hline
		$CO_2$ & 2349.91 & 140.7 & 3 & $v_1=0$,$v_2=0$, $i_2=1$, $v_3=1$, $J=1$, $r=1$ & 0.0949 & 0.69 \\
		$CO_2$ & 686.16 & 1.023 & 3 &  $v_1=0$,$v_2=1$, $i_2=1$, $J=1$, $r=1$  & 0.0949 & 0.69\\
		$CO_2$ & 3716.56 & 5.903 & 3 &  $v_1=1$,$v_2=1$, $i_2=0$, $J=1$, $r=1$ & 0.0949 & 0.69\\
		$H_2O$ & 1634.96 & 7.599 & 3 & $v_1=0$, $v_2=1$, $v_3=0$, $J=1$, $K_a=1$, $K_c=1$ & 0.1026 & 0.74 \\
		$H_2O$ & 3693.29 & 1.031 & 3 &  $v_1=1$, $v_2=0$, $v_3=0$, $J=1$, $K_a=1$, $K_c=1$ & 0.1049 & 0.71\\
		$H_2O$ & 3196.09 & 0.1982 & 3 &  $v_1=0$, $v_2=2$, $v_3=0$, $J=1$, $K_a=1$, $K_c=1$ & 0.1038 & 0.57\\
		$H_2O$ & 1601.34 & 14.79 & 3 &   $v_1=0$, $v_2=2$, $v_3=0$, $J=1$, $K_a=1$, $K_c=1$ & 0.0975 & 0.61\\
		\hline
	\end{tabular}
\end{table}	
Where the wave number $\tilde{\nu}$ is the frequency divided by the speed of light $c$
\begin{equation}
	\tilde{\nu} = \frac{1}{\lambda} = \frac{\nu}{c}
\end{equation}
  
\section{Required Equations}
  
The Intensity of radiation in the presence of an absorber is, \cite{Tokmakoff}
\begin{equation}
	I = I_0 e^{-N\alpha L}
\end{equation}
Where N is number density of the absorbing molecule, L the length of the absorbing region, and $\alpha$ the cross section of absorbtion.
Tokmakoff also gives for a single spectal line,
\begin{equation}
	\alpha = \frac{h\nu}{c} B_12
\end{equation}	
 Where $\nu$ is the frequency, and $B$ is the Einstein $B$ coefficient. HITRAN gives us the Einstein $A$ coefficent, which is related to $B$
 (at equilibrium) by \cite{Goldman}
 \begin{equation}
 	A_21 = \frac{8\pi h \nu^3}{c^3} B_21
 	\end{equation}
 and
 \begin{equation}
 	B_{12} = \frac{g_u}{g_l} B_{21}
 \end{equation}
 The ratio of lower to upper transition to the upper to lower transition is the ratio of the number of quantum states in the upper and lower
 states.
 So
 \begin{equation}
 	\alpha(\nu)\Delta_\nu = \frac{c^2}{8\pi\nu^2}g_u A_21
 \end{equation}
Of course we need the spectral lines for each absorber, labelled $M$, so
\begin{equation}
	I  = I_0 e^{-L\left(\sum_M N_m\ \sum_\nu \frac{c^2}{8\pi\Delta_\nu\nu^3} \right)}
\end{equation}

The input intensity is the Planck Radiation Law  of EM emission from a surface at temperature T, \cite{Brit1}

\begin{equation}
	I_0(\nu) = \frac{8\pi h \nu^5}{c^4}\frac{1}{e^\frac{-h\nu}{kT}} -1
\end{equation}

The width of the spectra line is commonly approximated as a Lorenzian, due to Natural and Collusional broadaning. \cite{KW} In fact the Lorenz shape estimate the tail of absorption, and
is often cut off at the edges, but we will not do that here. In HITRAN \cite{HITRANUnits} units, we lose the factor of $4\pi^2$ and the line width is proportional to the pressure $P$,
$\Gamma= P\gamma$, of the line width in table 1.

\begin{equation}
 \phi(\nu) = \frac{1}{\pi}\frac {\Gamma}{(\nu-\nu_0)^2 + \Gamma^2} 
\end{equation}

Then $\Delta_\nu$ can be approximately the line width $\Gamma$

HITRAN \cite{HITRANUnits} approximates the line width$\Gamma$ for air to 

\begin{equation}
	\Gamma(p, T) = \left(frac{T_{ref}}{T}\right)^{n_air} \gamma_{air} p]
\end{equation}

So now we have the absorption equation for surface temperature $T$ and Pressure $P$ at each height, so need a model of earths surface temperature and latitude $\theta$, and atmospheric 
pressure at height $h$, to complete our input equations. To begin with we will average the temperature over day and night, we need a rough input function of latitude that provides an average
surface air temperature, we find this Fuelner at al /cite{Feulner}, as graph Fig 5, we copy the data for Observations from 1961-1990 which we approcimate place the data points for each 10 degree latitude in the table below. We will then interpolate begin
the given points to provide a programmatic function to give a value for any latitude.

\begin{table}
	\caption{Reconstructed 1961-1990 Average Surface Air Temperature By Latitude}
	\centering
	\begin{tabular} { c c}
		\hline\hline
	    Latitude & Temperature Celsius \\ 
		\hline
			-90 & -20 \\
            -80 & -15 \\
            -70 & -8 \\
            -60 & -1 \\
            -50 & 10 \\
            -40 & 16 \\
            -30 & 20 \\
            -20 & 24 \\
            -10 & 26 \\
             0 & 27 \\
             10 & 27 \\
             20 & 25 \\
             30 & 20 \\
             40 & 17 \\
             50 & 7 \\
             60 & -1 \\
             70 & -8 \\
             80 & -15 \\
             90 & -17 \\
        \hline
    \end{tabular}
\end{table}    

Given the Pressure $P_0$ and Temperature $T_0$ at the Earths surface we can find the temperature, $T(h)$ and pressure a $P(h)$ a height $h$, using a Barometric formula, we use Lente and Osz's formule \cite(Lente)

\begin{equation}
	P(h) =  \left(P_0^\frac{1}{\nu+1}-\frac{Mg_0R_E P_0^\frac{1}{\nu+1}h}{(v+1)RT_0(R_E+h)}\right)^{\nu+1}
\end{equation} 
and
\begin{equation}
	T(h) = T_0 \left( 1 - \frac{Mg_0R_Eh}{(\nu+1)RT_0(R_E+h)}\right)
\end{equation} 
where $R$ is the gas constant, $R_E$ is the radius of the Earth, $m$ is the mass density of air $M=mN_A$ is the molar mass of air. $N_A$ IS Avogaro's constant, $g$ is acceleration due to gravity at the surface, $K$ is Boltzmann's constant,
$v$ is the number of degree of freedom of the average molecule in air $\nu\approx 5/2$, and $m$ is mass density of air. The formula works to about $30Km$, which is around 4 e-folds of exponentionly growth, as the
scale height of air pressure is around $7 Km$. 


\begin{thebibliography}{99}
	\bibitem{Mourya} Prof A. Mourya, Engineering Physics, Lecture Notes, \url{https://sites.google.com/site/puenggphysics/home/unit-i/relation-between-einstein-s-a-and-b-coefficient} (2015)
	\bibitem{Tokmakoff} Tokmakoff, Absoulute Cross Sections, Lecture Notes, \url{http://home.uchicago.edu/~tokmakoff/TDQMS/Notes/4.3._Abs_Cross-Sec_2-12-08.pdf}
    \bibitem{Brit1} Britannica, T. Editors of Encyclopaedia (Invalid Date). Planck's radiation law. Encyclopedia Britannica. \url{https://www.britannica.com/science/Plancks-radiation-law}
    \bibitem{KW} Kenneth Wood,University of St Andrews , Nebalae, Lecture 8, Line Widths \url{http://www-star.st-and.ac.uk/~kw25/teaching/nebulae/lecture08_linewidths.pdf}
    \bibitem{HITRANUnits} HITRAN Definitions and Units \url{https://hitran.org/docs/definitions-and-units/}
    \bibitem{Goldman} instein A-coefficients and statistical weights for molecular
    absorption transitions in the HITRAN database, Marie Simeckova, David Jacquemarta, Laurence S. Rothmana,
    Robert R. Gamacheb, Aaron Goldmanc. \url{https://lweb.cfa.harvard.edu/atmosphere/publications/2006-EinsteinA-JQSRT-98.pdf}
    \bibitem{Feulner} G. Feulner, S. Rahmstorf, A. Levermann, S. Volkwardt, On the Origin of Surface Air Temperature Difference between Hemispheres in the Earth's Present-Day Climate,
    pg 7136-7130 Vol 26 Jounral of Climate (2013)
	\bibitem{Lente} Lente, G., Ősz, K. Barometric formulas: various derivations and comparisons to environmentally relevant observations. ChemTexts 6, 13 (2020). https://doi.org/10.1007/s40828-020-0111-6
\end{thebibliography}

\end{document}
